Structured Genealogy Processes

### 2022-02-15 ms.Rnw
sgp probs update
updates in sgp thm 1 and thm2

\begin{figure}
  \resizebox{0.6\linewidth}{!}{
    \begin{tikzpicture}[scale=2.5]
      \usetikzlibrary{arrows}
      \tikzstyle{coordinate}=[inner sep=0pt,outer sep=0pt]
      \tikzstyle{det}=[color=black, thick, >=stealth]
      \tikzstyle{stoch}=[color=black, dashed, thick, >=stealth]
      \node (A) at (0,0) {$\Master_t$};
      \node (H) at (30:1) {$\Hist_t$};
      \node (X) at (30:2) {$\X_t$};
      \node (G) at (-30:1) {$\Gen_t$};
      \node (I) at (-30:2) {$\Inv_t$};
      \node (P) at (0:1.732) {$\Prune_t$};
      \node (O) at (1.732,.5) {$\Och_i$};
      \node (V) at (0:2.732) {$\Vis_t$};
      \node (W) at (2.732,.5) {$\Wch_i$};
      \draw [det,->] (A) -- (H);
      \draw [det,->] (H) -- (X);
      \draw [det,->] (A) -- (G);
      \draw [det,->] (G) -- (I) node[midway,below,sloped] {$\mathrm{ext}$};
      \draw [det,->] (G) -- (P) node[midway,below,sloped] {$\mathrm{prune}$};
      \draw [det,->] (P) -- (V) node[midway,below,sloped] {$\mathrm{obs}$};
      \draw [det,->] (P) -- (O);
      \draw [det,->] (V) -- (W);
      \draw [stoch,->] (H) -- (G);
    \end{tikzpicture}
  }
  \caption{
    \label{fig:constellation}
    Relations among the various Markov processes discussed in the paper.
    Deterministic maps are indicated with solid arrows;
    random maps are shown as dashed arrows.
    All the maps shown commute.
    $\X_t$ is the \emph{population process}, a model of the dynamics of some system, which we take as a starting point.
    $\Hist_t$ is the \emph{history process}, which records the full history of $\X_t$.
    $\Inv_t$ is the \emph{inventory process}: at each time $t$, $\Inv_t$ is an inventory of all extant individuals in the population, each of which has a globally unique name.
    $\Gen_t$ is the \emph{genealogy process}, which captures the precise genealogical relationships among all individuals in $\Inv_t$, as well as among any samples that have been taken from the population.
    $\Vis_t$ is the \emph{visible genealogy process}, which is $\Gen_t$ pruned so that only relationships among samples remain.
    Finally $\Wch_i$ is the \emph{embedded chain of the visible genealogy process}, which is $\Vis_{s_i}$, $s_i$ being the time of the $i$-th sample.
    All of these processes can be obtained via deterministic procedures applied to the \emph{master process} $\Master_t$, as described in the text.
  }
\end{figure}

\paragraph{Births}
Within a structured infective population, the parent and the child, at a given birth event, can be in different demes.
That is, an individual in Deme $i$ can give birth to a newborn individual in Deme $j$, where $i,j=1,\dots,J$. 
We then define the birth function $B_{ij}:\Uspace\to\N$ for any $i,j=1,\dots,J$. 
The set of birth events where an individual in Deme $i$ gives birth to $n$ individuals in Deme $j$ is $\Bset_{ij}\coloneqq\,B_{ij}^{-1}(\{1\})$. 
Trivially, the set of birth events in Deme $j$ is $\Bset_j\coloneqq\bigcup_{i}\Bset_{ij}$.

\paragraph{Migrations}
Infective demes are interchangeable and we define the migration function from Deme $i$ to $j$ for any $i\neq j$ and $i,j=1,\dots,J$ as $M_{ij}:\Uspace\to\N$ to represent the number of infectives migrating to Deme $j$ from Deme $i$.
Thus the set of migration is $\Mset_{ij}\coloneqq\,M^{-1}_{ij}(\{1\})$ since we assume single migration at a single migration event. 
Similarly, we disallow coninciding migration events and other events.

\paragraph{Samples}
We assume samples are taken serially and no births or deaths are coincided. 
We then define the sampling function on the probability space of the population process $\sg{X}_t$: $G:\Uspace\to\N_0$ is the number of samples taken at event $u$, then the set of sampling events is $\Sset\coloneqq\,G^{-1}(\{1\})$. 
Of course, the non-coincidence assumption can be easily relaxed while it's not the focus in this paper.

\paragraph{Deaths and population size}
The population, the set of infected individuals in the context of epidemiology, is structured, categorized into interchangeable demes. 
We can then define population size function and death functions in Deme $j$, respectively, as $I^{j}, D_{j}:\Xspace\to\N$ for any $j=1,\cdots,J$. 
\begin{equation}
  \jumprate_u(t,x) > 0 \Longrightarrow I^j(x+u)-I^j(x)=\sum_{i}\,B_{ij}(u)+\sum_{i}M_{ij}(u)-D_j(u)-\sum_{i}M_{ji}(u),
\end{equation}

for all $x, u\in\Xspace$. We disallow simultaneous death at a single death event at this stage and define the set of death events in Deme $j$ as $\Dset_j\coloneqq\,D^{-1}(\{1\})$. 
We also insist, birth and death events don't coincide for any demes, $i,j$, that is,  $\Bset_{j}\cap\Dset_{i}=\emptyset$. 
Eventually, we can have the function for total population size by letting $I(x)\coloneqq\sum_{j=1}^J\,I^j(x)$ and it's trivial that

\begin{equation}
  \jumprate_u(t,x) > 0 \Longrightarrow I(x+u)-I(x)=\sum_{i,j}B_{i,j}(u)-\sum_{j}D_j(u),
\end{equation}

@@

The above shows the rates of being infected, progression from exposed to infectious, recovery, and sampling, respectively. 
As in \citet{King2021}, the transmission $b(t)$ is time-dependent, and the sampling rate $\psi$ is any function, as long as the constraints in Section 2 are satisfied. 
Therefore, $I_1(\sg{X})=e$, $I_2(\sg{X})=i$, $I(\sg{X})=e+i$, $\Bset_{21}=\{\left(-1,1,0,0\right)\}$, $\Mset_{12}=\{\left(0,-1,1,0\right)\}$, $\Dset_2=\{\left(0,0,-1,0\right)\}$, $\Sset=\{\left(0,0,0,1\right)\}$, and $\Bset_{11}=\Bset_{12}=\Bset_{22}=\Mset_{21}=\Dset_1=\emptyset$.

@@

\QYL{the definition for sets of event times is problematic.}

Given a genealogy $\Gen$, the set of \emph{genealogical event times}, $\Eset(\Gen)$, is the set of all node times.
Its subsets, including those for different event types and those for different demes, are of interest.
In particular, we define
\begin{equation*}
  \begin{aligned}
    \Eset^j(\Gen) &\coloneqq \{t(\Node{p})\;|\;\Node{p}\in\Gen, z(\Node{p})\ \text{is}\ j\}, \\
    \Masterset^j(\Gen) &\coloneqq , \\
    \Cset^{j,j'}(\Gen) &\coloneqq , \\
    \Lset^j(\Gen) &\coloneqq , \\
    \Tset^j(\Gen) &\coloneqq , \\
    \Zset^{j,j'}(\Gen) &\coloneqq , \\
    \Qset^j(\Gen) &\coloneqq \Masterset^j(\Gen) \cap \Tset^j(\Gen).
  \end{aligned}
\end{equation*}
With these definitions,
$\Masterset^j(\Gen)$ comprises the internal node times of nodes that are derived from nodes in Deme $j$,
$\Cset^{j,j'}(\Gen)$ is the set of coalescent times of nodes in Deme $j$ and containing a ball in Deme $j'$,
$\Lset^j(\Gen)$ contains the leaf times where the tips are in Deme $j$,
$\Tset^j(\Gen)$ holds the sample times for samples in Deme $j$,
$\Zset^{j,j'}(\Gen)$ holds the deme-transition times derived from nodes in Deme $j$ and containing a ball in Deme $j'$,
and $\Qset^j(\Gen)$ is the set of \emph{direct-descent times} derived from samples in Deme $j$.

@@

\begin{equation}
  \ell^j(t,\Prune)\coloneqq
\end{equation}

@@

We define $\phi_{qk}^{j}\coloneqq\sum_{\omega\in\Omega_q}\dfrac{\indicator{\Inv_{t_{k-1}}^j(\omega)}(N_k(\omega))}{I^j(x_k-u_k)}$.

Note that $\ell_{1k}=0$ for all $k$.

@@

  % M
  \item There was a migration event, that is, $u_k\in\Mset^{j,j'}$ within $\halfopen{a_q,s_q}$, given that it was not presented in the pre-existing genealogy $\Och_{q-1}$, \ie $t_k\notin\Zset^{j,j'}(\Och_{q-1})$: 
    \begin{enumerate}
      \item if this migration event exactly involved the $q$-th sample and was presented as a deme-transition event in $\Och_q$, \ie $t_k\in\Zset^{j,j'}(\Och_q)$, then $\phi_{qk}^{j''}=\delta_{j',j''}/\left((I_{k-1}^{j}-\ell_{q,k-1}^{j})(I_{k}^{j'}-\ell_{qk}^{j'})\right)$;
      \item otherwise, $t_k\notin\Zset^{j,j'}(\Och_q)$, then $\phi_{qk}^{j''}=1-\delta_{j',j''}/\left((I_{k-1}^{j}-\ell_{q,k-1}^{j})(I_{k}^{j'}-\ell_{qk}^{j'})\right)$.
    \end{enumerate}

@@

To summarize, we have
\begin{equation}\label{eq:qcases}
  \phi_{qk}^{j''}=
  \begin{cases}
    1, &\text{if}\ t_k\notin\halfopen{a_q,s_q}\ \text{or}\ u_k\notin\Bset\cup\Sset\cup\Mset,\\

    % deme-transition time can evolve to coalescent time, when the new sample is the parent of a "cross-deme" birth
    1, &\text{if}\ t_k\in\Cset(\Och_{q-1})\cup\Qset(\Och_{q-1}),\\[12pt]
    \frac{\delta_{j',j''}}{(I_{k-1}^j-\ell_{q,k-1}^j)(I_k^{j'}-\ell_{qk}^{j'})}, &\text{if}\ t_k\in(a_q,s_q),\ t_k\in\Zset^{j,j'}(\Och_q),\ \text{and}\ u_k\in\Mset^{j,j'},\\[12pt]
    1-\frac{\delta_{j',j''}}{(I_{k-1}^j-\ell_{q,k-1}^{j})(I_k^{j'}-\ell_{qk}^{j'})}, &\text{if}\ t_k\in(a_q,s_q),\ t_k\notin\Zset^{j,j'}(\Och_q),\ \text{and}\ u_k\in\Mset^{j,j'},\\[12pt]

    % sample
    \frac{\delta_{j',j''}}{I_k^{j'}-\ell_{qk}^{j'}}, &\text{if}\ t_k=a_q\in\Lset^{j'}(\Och_{q-1})\ \text{and}\ u_k\in\Sset^{j'},\\[12pt]
    1-\frac{\delta_{j',j''}}{I_k^{j'}-\ell_{qk}^{j'}}, &\text{if}\ t_k\in(a_q,s_q)\setminus\Qset(\Och_{q-1})\ \text{and}\ u_k\in\Sset^{j'},\\[12pt]

@@

    % B, j!=j'
    % in Z
    \frac{\delta_{j,j''}}{I_k^{j}-\ell_{qk}^{j}}, &\text{if}\ t_k\in\Zset^{j,j'}(\Och_{q-1}),\ t_k=a_q\in\Cset^{j,j'}(\Och_q),\ u_k\in\Bset^{j,j'},\ \text{and}\ j\neq j',\\[12pt]
    1-\frac{\delta_{j,j''}}{I_k^{j}-\ell_{qk}^{j}}, &\text{if}\ t_k\in\Zset^{j,j'}(\Och_{q-1}),\ t_k\in\Zset^{j,j'}(\Och_q),\ u_k\in\Bset^{j,j'},\ \text{and}\ j\neq j',\\[12pt]
    % not in Z
    \frac{\delta_{j',j''}}{I_k^j(I_{k}^{j'}-\ell_{qk}^{j'})}, &\text{if}\ t_k\notin\Zset^{j,j'}(\Och_{q-1}),\ t_k\in\Zset^{j,j'}(\Och_q),\ u_k\in\Bset^{j,j'},\ \text{and}\ j\neq j',\\[12pt]
    \frac{\delta_{j',j''}}{I_k^j(I_{k}^{j'}-\ell_{qk}^{j'})}, &\text{if}\ t_k\notin\Zset^{j,j'}(\Och_{q-1}),\ t_k=a_q\in\Cset^{j,j'}(\Och_q),\ u_k\in\Bset^{j,j'},\ \text{and}\ j\neq j',\\[12pt]
    1-\frac{\delta_{j',j''}}{I_{k}^{j'}-\ell_{qk}^{j'}}, &\text{if}\ t_k\notin\Zset^{j,j'}(\Och_{q-1}),\ t_k\in(a_q,s_q),\ u_k\in\Bset^{j,j'},\ \text{and}\ j\neq j',\\[12pt]

@@

\begin{thm}\label{thm:prunehthm1}
  With the definitions as above, $P_{\Och_i|\Hist}(o|h) = \prod_{q=1}^{i}\prod_{k=1}^{K}\left(\sum_{j''=1}^{J}\phi_{qk}^{j''}\right)=\prod_{q=1}^{i}\prod_{k=1}^{K}\phi_{qk}^{j''_q}$, where $j''_q$ is the deme of the $q$-th sample lineage.
\end{thm}
We now exchange the order of multiplication over $q$ and $k$ then we can derive our \Cref{thm:prunethm2}.

\begin{thm}\label{thm:prunethm2}
  \begin{equation*}
    P_{\Prune_t|\Hist_t}(\Prune_t|h)=.
  \end{equation*}
\end{thm}
\begin{proof}
  Let $\Prune_t=\Och_i$ for some $i$.

  Last but not least, suppose $u_k\in\Bset^{j,j'}$.

>>>>

\begin{figure}
  \resizebox{0.6\linewidth}{!}{
    \begin{tikzpicture}[scale=1.8]
      \usetikzlibrary{arrows}
      \tikzstyle{coordinate}=[inner sep=0pt,outer sep=0pt]
      \tikzstyle{det}=[color=black, thick, >=stealth]
      \tikzstyle{stoch}=[color=black, dashed, thick, >=stealth]
      \node (A) at (0,0) {$\Master_t$};
      \node (H) at (30:1) {$\Hist_t$};
      \node (X) at (30:2) {$\X_t$};
      \node (G) at (-30:1) {$\Gen_t$};
      \node (I) at (-30:2) {$\Inv_t$};
      \node (P) at (0:1.732) {$\Prune_t$};
      \node (O) at (1.732,.5) {$\Och_i$};
      \node (V) at (0:2.732) {$\Vis_t$};
      \node (W) at (2.732,.5) {$\Wch_i$};
      \draw [det,->] (A) -- (H);
      \draw [det,->] (H) -- (X);
      \draw [det,->] (A) -- (G);
      \draw [det,->] (G) -- (I) node[midway,below,sloped] {$\mathrm{ext}$};
      \draw [det,->] (G) -- (P) node[midway,below,sloped] {$\mathrm{prune}$};
      \draw [det,->] (P) -- (V) node[midway,below,sloped] {$\mathrm{obs}$};
      \draw [det,->] (P) -- (O);
      \draw [det,->] (V) -- (W);
      \draw [stoch,->] (H) -- (G);
    \end{tikzpicture}
  }
  \caption{
    \label{fig:constellation}
    Relations among the various Markov processes discussed in the paper.
    Deterministic maps are indicated with solid arrows;
    random maps are shown as dashed arrows.
    All the maps shown commute.
    $\X_t$ is the \emph{population process}, a model of the dynamics of some system, which we take as a starting point.
    $\Hist_t$ is the \emph{history process}, which records the full history of $\X_t$.
    $\Inv_t$ is the \emph{inventory process}: at each time $t$, $\Inv_t$ is an inventory of all extant individuals in the population, each of which has a globally unique name.
    $\Gen_t$ is the \emph{genealogy process}, which captures the precise genealogical relationships among all individuals in $\Inv_t$, as well as among any samples that have been taken from the population.
    $\Vis_t$ is the \emph{visible genealogy process}, which is $\Gen_t$ pruned so that only relationships among samples remain.
    Finally $\Wch_i$ is the \emph{embedded chain of the visible genealogy process}, which is $\Vis_{s_i}$, $s_i$ being the time of the $i$-th sample.
    All of these processes can be obtained via deterministic procedures applied to the \emph{master process} $\Master_t$, as described in the text.
  }
\end{figure}

\paragraph{Births}
Within a structured infective population, the parent and the child, at a given birth event, can be in different demes.
That is, an individual in Deme $j$ can give birth to a newborn individual in Deme $j'$, where $j,j'=1,\dots,J$. 
We then define the birth function $B_{j,j'}:\Uspace\to\N$ for any $j,j'=1,\dots,J$. 
The set of birth events where an individual in Deme $j$ gives birth to $n$ individuals in Deme $j'$ is $\Bset^{j,j'}\coloneqq\,B_{j,j'}^{-1}(\{1\})$. 
Trivially, the set of birth events in Deme $j'$ is $\Bset^{j'}\coloneqq\bigcup_{j}\Bset^{j,j}$.

\paragraph{Migrations}
Infective demes are interchangeable and we define the migration function from Deme $j$ to $j'$ for any $j\neq j'$ and $j,j'=1,\dots,J$ as $M_{j,j'}:\Uspace\to\N$ to represent the number of infectives migrating to Deme $j'$ from Deme $j$.
Thus the set of migration is $\Mset^{j,j'}\coloneqq\,M^{-1}_{j,j'}(\{1\})$ since we assume single migration at a single migration event. 
Similarly, we disallow coninciding migration events and other events.

\paragraph{Samples}
We assume samples are taken serially and no births or deaths are coincided. 
We then define the sampling function on the probability space of the population process $\sg{X}_t$: $G_j:\Uspace\to\N_0$ is the number of samples from Deme $j$ taken at event $u$, then the set of sampling events is $\Sset^j\coloneqq\,G_j^{-1}(\{1\})$. 
Of course, the non-coincidence assumption can be easily relaxed while it's not the focus in this paper.

\paragraph{Deaths and population size}
The population, the set of infected individuals in the context of epidemiology, is structured, categorized into interchangeable demes. 
We can then define population size function and death functions in Deme $j$, respectively, as $I^{j}, D_{j}:\Xspace\to\N$ for any $j=1,\cdots,J$. 
\begin{equation}
  \jumprate_u(t,x) > 0 \Longrightarrow I^j(x+u)-I^j(x)=\sum_{j'}\,B_{j',j}(u)+\sum_{j'}M_{j',j}(u)-D_j(u)-\sum_{j'}M_{j,j'}(u),
\end{equation}

for all $x, u\in\Xspace$. We disallow simultaneous death at a single death event at this stage and define the set of death events in Deme $j$ as $\Dset^j\coloneqq\,D_j^{-1}(\{1\})$. 
We also insist, birth and death events don't coincide for any demes, $j,j'$, that is,  $\Bset_{j}\cap\Dset_{j'}=\emptyset$. 
Eventually, we can have the function for total population size by letting $I(x)\coloneqq\sum_{j=1}^J\,I^j(x)$ and it's trivial that

\begin{equation}
  \jumprate_u(t,x) > 0 \Longrightarrow I(x+u)-I(x)=\sum_{j,j'}B_{j,j'}(u)-\sum_{j}D_j(u),
\end{equation}

@@

The above shows the rates of being infected, progression from exposed to infectious, recovery, and sampling, respectively. 
As in \citet{King2021}, the transmission $b(t)$ is time-dependent, and the sampling rate $\psi$ is any function, as long as the constraints in Section 2 are satisfied. 
Therefore, $I_1(\sg{X})=e$, $I_2(\sg{X})=i$, $I(\sg{X})=e+i$, $\Bset^{2,1}=\{\left(-1,1,0,0\right)\}$, $\Mset^{1,2}=\{\left(0,-1,1,0\right)\}$, $\Dset^2=\{\left(0,0,-1,0\right)\}$, $\Sset^2=\{\left(0,0,0,1\right)\}$, and $\Bset^{1,1}=\Bset^{1,2}=\Bset^{2,2}=\Mset^{2,1}=\Dset^1=\Sset^1=\emptyset$.

@@

% \QYL{the definition for sets of event times is problematic.}

Given a genealogy $\Gen$, the set of \emph{genealogical event times}, $\Eset(\Gen)$, is the set of all node times.
Its subsets, including those for different event types and those for different demes, are of interest.
In particular, we define
\begin{equation*}
  \begin{aligned}
    \Eset^j(\Gen) &\coloneqq \{t(\Node{p})\;|\;\Node{p}\in\Gen, z(\Node{p})=j\}, \\
    \Masterset^j(\Gen) &\coloneqq \{t(\Node{p})\;|\;\Node{p}\in\Gen, z(\Node{p})=j, w(\Node{p})\ \text{contains one green ball but no purple balls}\}, \\
    \Cset^{j,j'}(\Gen) &\coloneqq \{t(\Node{p})\;|\;\Node{p}\in\Gen, z(\Node{p})=j, w(\Node{p})\ \text{contains two green balls, one from Deme $j$ and the other from $j'$}\}, \\
    \Lset^j(\Gen) &\coloneqq \{t(\Node{p})\;|\;\Node{p}\in\Gen, z(\Node{p})=j, w(\Node{p})\ \text{contains no green balls}\}, \\
    \Tset^j(\Gen) &\coloneqq \{t(\Node{p})\;|\;\Node{p}\in\Gen, z(\Node{p})=j, w(\Node{p})\ \text{contains a blue ball}\}, \\
    \Zset^{j,j'}(\Gen) &\coloneqq \{t(\Node{p})\;|\;\Node{p}\in\Gen, z(\Node{p})=j, w(\Node{p})\ \text{contains a purple ball and a green ball from Deme $j'$}\}, \\
    \Qset^j(\Gen) &\coloneqq \Masterset^j(\Gen) \cap \Tset^j(\Gen).
  \end{aligned}
\end{equation*}
With these definitions,
$\Masterset^j(\Gen)$ comprises the internal node times of nodes that are derived from nodes in Deme $j$,
$\Cset^{j,j'}(\Gen)$ is the set of coalescent times of nodes in Deme $j$ and containing a ball in Deme $j'$,
$\Lset^j(\Gen)$ contains the leaf times where the tips are in Deme $j$,
$\Tset^j(\Gen)$ holds the sample times for samples in Deme $j$,
$\Zset^{j,j'}(\Gen)$ holds the deme-transition times derived from nodes in Deme $j$ and containing a ball in Deme $j'$,
and $\Qset^j(\Gen)$ is the set of \emph{direct-descent times} derived from samples in Deme $j$.

@@

\begin{align}
\begin{split}
  \ell^j(t,\Prune) &\coloneqq \sum_{j'}\sum_{e\in\Cset^{j',j}(\Prune)} \indicator{\halfopen{e,\infty}}(t)\;-\;\sum_{e\in\Lset^j(\Prune)} \indicator{\halfopen{e,\infty}}(t) \\
  +&\;\sum_{j'}\left(\sum_{e\in\Zset^{j',j}(\Prune)}\indicator{\halfopen{e,\infty}}(t)\;-\;\sum_{e\in\Zset^{j,j'}(\Prune)}\indicator{\halfopen{e,\infty}}(t)\right).
\end{split}
\end{align}

@@

We define $\phi_{qk}^{j}\coloneqq\sum_{\omega\in\Omega_q}\dfrac{\indicator{\Inv_{t_{k-1}}^j(\omega)}(N_k(\omega))}{I^j(x_k-u_k)}$, 
and let $I_k^j=I^j(x_k)$, $\ell_{qk}^{j}=\ell^j(t_k, \Och_{q-1})$.
Note that $\ell^j_{1k}=0$ for all $k$ and $j$.

@@

  % M
  \item There was a migration event, that is, $u_k\in\Mset^{j,j'}$ within $\halfopen{a_q,s_q}$, given that it was not presented in the pre-existing genealogy $\Och_{q-1}$, \ie $t_k\notin\Zset^{j,j'}(\Och_{q-1})$: 
    \begin{enumerate}
      \item if this migration event had been presented in previous genealogy $\Och_{q-1}$, \ie $t_k\in\Zset^{j,j'}(\Och_{q-1})$, since no simultaneous migration events occurred at the same time and $t_k\in\Zset^{j,j'}{\Och_q}$ for sure, $\phi_{qk}^{j''}=1$;
      \item if this migration event exactly involved the $q$-th sample and was presented as a deme-transition event in $\Och_q$, \ie $t_k\in\Zset^{j,j'}(\Och_q)$, then $\phi_{qk}^{j''}=\delta_{j',j''}/\left((I_{k}^{j}-\ell_{q,k}^{j})(I_{k}^{j'}-\ell_{qk}^{j'})\right)$;
      \item otherwise, $t_k\notin\Zset^{j,j'}(\Och_q)$, then $\phi_{qk}^{j''}=1-\delta_{j',j''}/\left((I_{k}^{j}-\ell_{q,k}^{j})(I_{k}^{j'}-\ell_{qk}^{j'})\right)$.
    \end{enumerate}

@@

To summarize, we have
\begin{equation}\label{eq:qcases}
  \phi_{qk}^{j''}=
  \begin{cases}
    1, &\text{if}\ t_k\notin\halfopen{a_q,s_q}\ \text{or}\ u_k\notin\Bset\cup\Sset\cup\Mset,\\
    1, &\text{if}\ t_k\in\Cset(\Och_{q-1})\cup\Qset(\Och_{q-1}),\\[12pt]

    % M
    1, &\text{if}\ t_k\in(a_q,s_q),\ t_k\in\Zset^{j,j'}(\Och_{q-1}),\ t_k\in\Zset^{j,j'}(\Och_{q}),\ \text{and}\ u_k\in\Mset^{j,j'},\\[12pt]
    \frac{\delta_{j',j''}}{(I_{k}^j-\ell_{q,k}^j)(I_k^{j'}-\ell_{qk}^{j'})}, &\text{if}\ t_k\in(a_q,s_q),\ t_k\in\Zset^{j,j'}(\Och_q),\ \text{and}\ u_k\in\Mset^{j,j'},\\[12pt]
    1-\frac{\delta_{j',j''}}{(I_{k}^j-\ell_{q,k}^{j})(I_k^{j'}-\ell_{qk}^{j'})}, &\text{if}\ t_k\in(a_q,s_q),\ t_k\notin\Zset^{j,j'}(\Och_q),\ \text{and}\ u_k\in\Mset^{j,j'},\\[12pt]

    % G
    \frac{\delta_{j',j''}}{I_k^{j'}-\ell_{qk}^{j'}}, &\text{if}\ t_k=a_q\in\Lset^{j'}(\Och_{q-1})\ \text{and}\ u_k\in\Sset^{j'},\\[12pt]
    1-\frac{\delta_{j',j''}}{I_k^{j'}-\ell_{qk}^{j'}}, &\text{if}\ t_k\in(a_q,s_q)\setminus\Qset(\Och_{q-1})\ \text{and}\ u_k\in\Sset^{j'},\\[12pt]

@@

    % B, j!=j'
    % in Z
    \frac{\delta_{j,j''}}{I_k^{j}-\ell_{qk}^{j}}, &\text{if}\ t_k\in\Zset^{j,j'}(\Och_{q-1}),\ t_k=a_q\in\Cset^{j,j'}(\Och_q),\ u_k\in\Bset^{j,j'},\ \text{and}\ j\neq j',\\[12pt]
    1-\frac{\delta_{j,j''}}{I_k^{j}-\ell_{qk}^{j}}, &\text{if}\ t_k\in(a_q,s_q),\ t_k\in\Zset^{j,j'}(\Och_{q-1}),\ t_k\in\Zset^{j,j'}(\Och_q),\ u_k\in\Bset^{j,j'},\ \text{and}\ j\neq j',\\[12pt]
    % not in Z
    \frac{\delta_{j',j''}}{I_k^j(I_{k}^{j'}-\ell_{qk}^{j'})}, &\text{if}\ t_k\in(a_q,s_q),\ t_k\notin\Zset^{j,j'}(\Och_{q-1}),\ t_k\in\Zset^{j,j'}(\Och_q),\ u_k\in\Bset^{j,j'},\ \text{and}\ j\neq j',\\[12pt]
    \frac{\delta_{j',j''}}{I_k^j(I_{k}^{j'}-\ell_{qk}^{j'})}, &\text{if}\ t_k\notin\Zset^{j,j'}(\Och_{q-1}),\ t_k=a_q\in\Cset^{j,j'}(\Och_q),\ u_k\in\Bset^{j,j'},\ \text{and}\ j\neq j',\\[12pt]
    1-\frac{\delta_{j',j''}}{I_{k}^{j'}-\ell_{qk}^{j'}}, &\text{if}\ t_k\in(a_q,s_q),\ t_k\notin\Zset^{j,j'}(\Och_{q-1}),\ u_k\in\Bset^{j,j'},\ \text{and}\ j\neq j',\\[12pt]

@@

\begin{thm}\label{thm:prunehthm1}
  With the definitions as above, $P_{\Och_i|\Hist}(o|h) = \prod_{q=1}^{i}\prod_{k=1}^{K}\left(\sum_{j''=1}^{J}\phi_{qk}^{j''}\right)=\prod_{q=1}^{i}\prod_{k=1}^{K}\phi_{qk}^{j''_{qk}}$, where $j''_{qk}$ is the deme of the $q$-th sample lineage at time $t_k$.
\end{thm}
We now exchange the order of multiplication over $q$ and $k$ then we can derive our \Cref{thm:prunethm2}.

\begin{thm}\label{thm:prunethm2}
  \begin{equation*}
    P_{\Prune_t|\Hist_t}(\Prune_t|h)=.
  \end{equation*}
\end{thm}
\begin{proof}
  Let $\Prune_t=\Och_i$ for some $i$. We define $\Phi_{k}\coloneqq\prod_{q=1}^{i}\phi_{qk}^{j''_{qk}}$ for each $k$ and as above, $I_k^{j}=I_j(x_k)$, $\ell_{qk}^{j}=\ell^j(t,\Och_{q-1})$. For the first two cases in \Cref{eq:qcases}, $\Phi_k=1$.

  Consider $u_k\in\Mset^{j,j'}$. Let $q^*=\{q\;\vert\;t_k=\Zset^{j,j'}(\Och_q)\setminus\Zset^{j,j'}(\Och_{q-1})\}$ and $\ell^*=\ell_{q^*k}^{j''_{q^*k}}$. We then define $Q_k=\{q\;\vert\;j''_{qk}=j'\ \text{and}\ q<q^*\}$ as an ordered set and denote $[Q_k]_m$ as the $m$-th element in $Q_k$, where, $m=1,\dots, \ell^*$. 
  \begin{enumerate}[(a)]
    \item If $q<q^*$, then $t_k\notin\Zset^{j,j'}(\Och_q)$. We will have
      \begin{equation*}
        \Phi_k = \prod_{q\in Q_k} \left(1-\frac{1}{(I_k^j-\ell_{qk}^j)(I_k^{j'}-\ell_{qk}^{j'})}\right).
      \end{equation*}
      Note that, $\{\ell_{qk}^{j'}\;\vert\;q\in Q_k\}=\{0,1,\dots,\ell^*-1\}$ then the previous equation becomes
      \begin{equation*}
        \Phi_k = \prod_{m=1}^{\ell*} \left(1-\frac{1}{\mathcal{Y}_m (I_k^{j'}-m+1)}\right) = \prod_{m=1}^{\ell*} \frac{I_k^{j'}-m+1-(\mathcal{Y}_m)^{-1}}{I_k^{j'}-m+1},
      \end{equation*}
      where $\mathcal{Y}_m\coloneqq I_k^{j}-\ell_{[Q_k]_m k}^{j}$.
  \end{enumerate}
